---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
Brain-Computer Interaction:

```{r}
library(caret)
library(lattice)
library(ggplot2)
library(RSNNS)
library(kernlab)


library (e1071)
library(MASS)
library (class)
library(nnet)
library(randomForest)

set.seed(1111)
```

##Section 1: Preprocess data

Functions to preprocess the data
```{r}
subject.nzv <- function(subject.predictors)
{
  nzv <- nearZeroVar(subject.predictors)
  print(paste0("Variables that their variance are almost zero: ", nzv))
  if(length(nzv) > 0){
    subject.predictors <- subject.predictors[, -nzv]
  }
  return(subject.predictors)
  
}

subject.highlyCor <- function(subject.predictors)
{
  descrCor <-  cor(subject.predictors)
  highlyCorDescr <- findCorrelation(descrCor, cutoff = .95)
  
  print(sort(highlyCorDescr))
  print(paste0("Variables that that are highly correlated: ", highlyCorDescr))
  if(length(highlyCorDescr) > 0)
  {
    subject.predictors <- subject.predictors[,-highlyCorDescr]
  }
  return(subject.predictors)
  
}

subject.linearCombos <- function(subject.predictors)
{
  comboInfo <- findLinearCombos(subject.predictors)
  print(paste0("columns positions can be removed to eliminate the linear dependencies: ",comboInfo))
  if(! is.null(comboInfo$remove)){
   subject.predictors <- subject.predictors[, -comboInfo$remove]
  }

  return(subject.predictors)
  
}

subject.preprocess<- function(subject.predictors)
  
{
  subject.predictors<-subject.nzv(subject.predictors)
  subject.predictors<-subject.highlyCor(subject.predictors)
  subject.predictors<-subject.linearCombos(subject.predictors)
  return(subject.predictors)
}



subject.nzv.return <- function(subject.predictors)
{
  nzv <- nearZeroVar(subject.predictors)
  print(paste("Variables that their variance are almost zero: ", nzv))
  return(nzv)
  
}

subject.highlyCor.return <- function(subject.predictors,cutoff)
{
  descrCor <-  cor(subject.predictors)
  highlyCorDescr <- findCorrelation(descrCor, cutoff = cutoff)
  print(paste("Variables that that are highly correlated: ", highlyCorDescr))
  return(highlyCorDescr)
  
}

subject.linearCombos.return <- function(subject.predictors)
{
  comboInfo <- findLinearCombos(subject.predictors)
  print(paste("columns positions can be removed to eliminate the linear dependencies: ",comboInfo))
  return(comboInfo$remove)
  
}

subject.preprocess.remove<- function(subject.predictors,cutoff)
  
{
  
  remove<-subject.nzv.return(subject.predictors)
  remove<-c(remove,subject.highlyCor.return(subject.predictors,cutoff))
  remove<-c(remove,subject.linearCombos.return(subject.predictors))
  return(sort(remove))
}



refactor <-function(subject_psa){
  
  output <- matrix(ncol=9, nrow=dim(subject_psa)[1])

  x <- 1
  eegData <- c('C3', 'Cz', 'C4', 'CP1', 'CP2', 'P3', 'Pz', 'P4','Target')
  newVector <- data.frame()
  
  for (id in 1:8)
  {
    y <- x+11
    output[,id] <- rowMeans(subject_psa[,x:y])
    x<-y+1
    
    
  }
  output[,9] <- subject_psa[,97]
  output <- data.frame(output)
  colnames(output) <- make.names(eegData)
  return(output)
}

```


Preprocess for psd data:
```{r}
subject1_1psd <- read.table("data_psd/train_subject1_psd01.asc", header = FALSE,  dec = ".")
subject1_2psd <- read.table("data_psd/train_subject1_psd02.asc", header = FALSE,  dec = ".")
subject1_3psd <- read.table("data_psd/train_subject1_psd03.asc", header = FALSE,  dec = ".")

subject1_psd <- rbind(subject1_1psd,subject1_2psd)
subject1_psd <- rbind(subject1_psd,subject1_3psd)
subject1_psd.refactor <-refactor(subject1_psd)


subject2_1psd <- read.table("data_psd/train_subject2_psd01.asc", header = FALSE,  dec = ".")
subject2_2psd <- read.table("data_psd/train_subject2_psd02.asc", header = FALSE,  dec = ".")
subject2_3psd <- read.table("data_psd/train_subject2_psd03.asc", header = FALSE,  dec = ".")

subject2_psd <- rbind(subject2_1psd,subject2_2psd)
subject2_psd <- rbind(subject2_psd,subject2_3psd)
subject2_psd.refactor <-refactor(subject2_psd)

subject3_1psd <- read.table("data_psd/train_subject3_psd01.asc", header = FALSE,  dec = ".")
subject3_2psd <- read.table("data_psd/train_subject3_psd02.asc", header = FALSE,  dec = ".")
subject3_3psd <- read.table("data_psd/train_subject3_psd03.asc", header = FALSE,  dec = ".")


subject3_psd <- rbind(subject3_1psd,subject3_2psd)
subject3_psd <- rbind(subject3_psd,subject3_3psd)
subject3_psd.refactor <-refactor(subject3_psd)

maxVarBefore <- dim(subject1_psd.refactor)[2]
cutoff <- 0.7
subject1_psd.variablesDelete<-subject.preprocess.remove(subject1_psd.refactor[,1:(maxVarBefore-1)],cutoff)

if(length(subject1_psd.variablesDelete)>0){
  subject1_psd.refactor <- subject1_psd.refactor[,-subject1_psd.variablesDelete]
  subject2_psd.refactor <- subject2_psd.refactor[,-subject1_psd.variablesDelete]
  subject3_psd.refactor <- subject3_psd.refactor[,-subject1_psd.variablesDelete]
}
maxVarBefore <- dim(subject1_psd.refactor)[2]
subject2_psd.variablesDelete<-subject.preprocess.remove(subject2_psd.refactor[,1:(maxVarBefore-1)],cutoff)
if(length(subject2_psd.variablesDelete)>0){
subject1_psd.refactor <- subject1_psd.refactor[,-subject2_psd.variablesDelete]
subject2_psd.refactor <- subject2_psd.refactor[,-subject2_psd.variablesDelete]
subject3_psd.refactor <- subject3_psd.refactor[,-subject2_psd.variablesDelete]
}
maxVarBefore <- dim(subject3_psd.refactor)[2]
subject3_psd.variablesDelete<-subject.preprocess.remove(subject3_psd.refactor[,1:(maxVarBefore-1)],cutoff)

if(length(subject3_psd.variablesDelete)>0){
subject1_psd.refactor <- subject1_psd.refactor[,-subject3_psd.variablesDelete]
subject2_psd.refactor <- subject2_psd.refactor[,-subject3_psd.variablesDelete]
subject3_psd.refactor <- subject3_psd.refactor[,-subject3_psd.variablesDelete]
}

print(paste("subject  1: ",subject1_psd.variablesDelete ))
print(paste("subject  2: ",subject2_psd.variablesDelete ))
print(paste("subject  3: ",subject3_psd.variablesDelete ))


```
Preprocess for  row data:
```{r}

subject1_1raw <- read.table("subject1/train_subject1_raw01.asc", header = FALSE,  dec = ".")
subject1_2raw <- read.table("subject1/train_subject1_raw02.asc", header = FALSE,  dec = ".")
subject1_3raw <- read.table("subject1/train_subject1_raw03.asc", header = FALSE,  dec = ".")


subject1_raw <- rbind(subject1_1raw,subject1_2raw)
subject1_raw <- rbind(subject1_raw,subject1_3raw)


subject2_1raw <- read.table("subject2/train_subject2_raw01.asc", header = FALSE,  dec = ".")
subject2_2raw <- read.table("subject2/train_subject2_raw02.asc", header = FALSE,  dec = ".")
subject2_3raw <- read.table("subject2/train_subject2_raw03.asc", header = FALSE,  dec = ".")


subject2_raw <- rbind(subject2_1raw,subject2_2raw)
subject2_raw <- rbind(subject2_raw,subject2_3raw)

subject3_1raw <- read.table("subject3/train_subject3_raw01.asc", header = FALSE,  dec = ".")
subject3_2raw <- read.table("subject3/train_subject3_raw02.asc", header = FALSE,  dec = ".")
subject3_3raw <- read.table("subject3/train_subject3_raw03.asc", header = FALSE,  dec = ".")


subject3_raw <- rbind(subject3_1raw,subject3_2raw)
subject3_raw <- rbind(subject3_raw,subject3_3raw)

maxVarBefore <- dim(subject3_raw)[2]
cutoff <- 0.98
subject1_raw.variablesDelete<-subject.preprocess.remove(subject1_raw[,1:(maxVarBefore-1)],cutoff)

if(length(subject1_raw.variablesDelete)>0){
subject1_raw <- subject1_raw[,-subject1_raw.variablesDelete]
subject2_raw <- subject2_raw[,-subject1_raw.variablesDelete]
subject3_raw <- subject3_raw[,-subject1_raw.variablesDelete]
}
maxVarBefore <- dim(subject3_raw)[2]
subject2_raw.variablesDelete<-subject.preprocess.remove(subject2_raw[,1:(maxVarBefore-1)],cutoff)

if(length(subject2_raw.variablesDelete)>0){
subject1_raw <- subject1_raw[,-subject2_raw.variablesDelete]
subject2_raw <- subject2_raw[,-subject2_raw.variablesDelete]
subject3_raw <- subject3_raw[,-subject2_raw.variablesDelete]
}
maxVarBefore <- dim(subject3_raw)[2]

subject3_raw.variablesDelete<-subject.preprocess.remove(subject3_raw[,1:(maxVarBefore-1)],cutoff)

if(length(subject3_raw.variablesDelete)>0){
subject1_raw <- subject1_raw[,-subject3_raw.variablesDelete]
subject2_raw <- subject2_raw[,-subject3_raw.variablesDelete]
subject3_raw <- subject3_raw[,-subject3_raw.variablesDelete]
}

print(paste("subject  1: ",subject1_raw.variablesDelete ))
print(paste("subject  2: ",subject2_raw.variablesDelete ))
print(paste("subject  3: ",subject3_raw.variablesDelete ))



```

##Section 2: Linear methods

Linear Functions: (LDA,QDA,Naive Bayes)
```{r}
subject.10foldCV.linear <- function(yourData,numClassify,maxVar)
{
  print("start foldCV")
  yourData<-yourData[sample(nrow(yourData)),]
  
  folds <- cut(seq(1,nrow(yourData)),breaks=10,labels=FALSE)
  errorRate <-0
  #Perform 10 fold cross validation
  for(i in 1:10){
      testIndexes <- which(folds==i,arr.ind=TRUE)
      newErrorRate <- subject.get.iteration.linear(numClassify,yourData,maxVar,testIndexes)
      errorRate <- errorRate + newErrorRate
  }
  print("finish foldCV")
  return(errorRate/10)
}


subject.TrainTest.linear <- function(yourData,numClassify,maxVar)
{
  yourData<-yourData[sample(nrow(yourData)),]
  subject.learn <- subject.get.learner(yourData,maxVar)
  errorRate <- subject.get.iteration.linear(numClassify,yourData,maxVar,subject.learn)
  return(errorRate)
}
subject.get.iteration.linear <- function(number,subject,maxVar,subject.learn)
{
  dataTrain <- subject[subject.learn,]
  classifier <- subject.get.classifier.linear(number,dataTrain,maxVar)
  errorRate <- subject.errorRate(number,subject,classifier,maxVar,subject.learn)
  return(errorRate)
}


subject.errorRate <- function(number,subject,subject.model,maxVar,subject.learn){
  ntrain <- length(subject.learn)
  ntest <-  dim(subject[-subject.learn,])[1]

    subject.predTest <- predict(subject.model, subject[-subject.learn,1:maxVar])
    if(number ==3 ){subject.resultsTest <- table(subject.predTest, subject[-subject.learn,(maxVar+1)]) }
    else{subject.resultsTest <- table(subject.predTest$class, subject[-subject.learn,(maxVar+1)]) }
  
  subject.errorRate.Test <- 1-(sum(diag(subject.resultsTest))/ntest)
  return(subject.errorRate.Test)
  
}

subject.get.classifier.linear<- function(number,subject,maxVar)
{
if(number == 1){method<- subject.get.lda(subject,maxVar)}
  else if(number == 2){method<- subject.get.qda(subject,maxVar)}
  else if(number==3){method<- subject.get.naiveBayes(subject,maxVar)} 
  
  return(method)
}


subject.get.message.method.linear<- function(number)
{
if(number == 1){message<- "LDA "}
  else if(number == 2){message<- " QDA "}
  else if(number==3){message<-" Naive Bayes"} 
  
  return(message)
}



subject.get.lda <- function(subject,maxVar)
{
  subject.model <- lda (x=subject[,1:maxVar], grouping=subject[,(maxVar+1)])
  return(subject.model)
} 

subject.get.qda <- function(subject,maxVar)
{
  subject.model <- qda (x=subject[,1:maxVar], grouping=subject[,(maxVar+1)])
  return(subject.model)
}

subject.get.naiveBayes <- function(subject,maxVar)
{
  subject.model <- naiveBayes(subject[,(maxVar+1)] ~ ., data =subject[,1:maxVar])
  return(subject.model)
}
subject.get.learner <- function(subject,maxVar)
{
  
  N <- nrow(subject)
  subject.learn <- sample(1:N, round(2*N/3))
  return(subject.learn)
}

subject.linearMethod <- function(subject,nameSubject)
{
 
  maxVar <- dim(subject)[2]-1
  subject[,(maxVar+1)] <- as.factor(subject[,(maxVar+1)])
    
    for( numberMethod in 1:3){
     
     messageMethod <-subject.get.message.method.linear(numberMethod)
     subject.error.10foldCV<- subject.10foldCV.linear (subject,numberMethod,maxVar)
     subject.error.TrainTest <- subject.TrainTest.linear(subject,numberMethod,maxVar)
     
     message <- paste("subject ", nameSubject , "using classifier ", messageMethod, " with 10foldCV ", subject.error.10foldCV, " with TrainTest resampling ", subject.error.TrainTest)
     print(message)

    }
  

}


```
Linear methods for all the subjects
```{r}

subject.linearMethod(subject1_psd.refactor,"1 with psd data")
subject.linearMethod(subject2_psd.refactor,"2 with psd data")
subject.linearMethod(subject3_psd.refactor,"3 with psd data")

subject.linearMethod(subject1_raw,"1 with raw data")
subject.linearMethod(subject2_raw,"2 with raw data")
subject.linearMethod(subject3_raw,"3 with raw data")
```

KNN functions:

```{r}

subject.get.knn <- function(subject,maxVar,learn,k)
{
  subject.model <-  knn (subject[learn,1:maxVar],subject[-learn,1:maxVar] ,subject[learn,(maxVar+1)],k=k,prob=TRUE)
  return(subject.model)
}



subject.errorRate.KNN <- function(subject,subject.model,maxVar,subject.learn){
  ntrain <- length(subject.learn)
  ntest <-  dim(subject[-subject.learn,])[1]

   subject.resultsTest <- table(subject.model, subject[-subject.learn,(maxVar+1)]) 
  subject.errorRate.Test <- 1-(sum(diag(subject.resultsTest))/ntest)
  return(subject.errorRate.Test)
  
}


subject.get.iteration.knn <- function(subject,maxVar,subject.learn,k)
{
  classifier <- subject.get.knn(subject,maxVar,subject.learn,k)
  errorRate <- subject.errorRate.KNN(subject,classifier,maxVar,subject.learn)
  return(errorRate)
}


subject.10foldCV.knn <- function(yourData,maxVar,k)
{
  print("start foldCV")
  yourData<-yourData[sample(nrow(yourData)),]
  
  folds <- cut(seq(1,nrow(yourData)),breaks=10,labels=FALSE)
  errorRate <-0
  #Perform 10 fold cross validation
  for(i in 1:10){
      testIndexes <- which(folds==i,arr.ind=TRUE)
      newErrorRate <- subject.get.iteration.knn(yourData,maxVar,testIndexes,k)
      errorRate <- errorRate + newErrorRate
  }
  print("finish foldCV")
  return(errorRate/10)
}


subject.TrainTest.knn <- function(yourData,maxVar,k)
{
  yourData<-yourData[sample(nrow(yourData)),]
  subject.learn <- subject.get.learner(yourData,maxVar)
  errorRate <- subject.get.iteration.knn(yourData,maxVar,subject.learn,k)
  return(errorRate)
}

subject.findk.KNN <- function(trainData,maxVar,learner)
{
  leastK<-0
  errorRate <- 10000
  for(k in seq(1,20,3))
  {
    
    newErrorRate <- subject.10foldCV.knn(trainData,maxVar,k)
    if(newErrorRate < errorRate)
    {
      errorRate <- newErrorRate
      leastK <- k
    }
    print(paste0("iteration k: ",k," errorRate ", errorRate))
    
  }
  return(leastK)
  
}


subject.KNNMethod <- function(subject,nameSubject)
{
 
  maxVar <- dim(subject)[2]-1
  subject[,(maxVar+1)] <- as.factor(subject[,(maxVar+1)])
  learner <- subject.get.learner(subject,maxVar)
  trainData <- subject[learner,]
  k<-subject.findk.KNN(trainData,maxVar,learner)
  errorRate <- subject.get.iteration.knn(subject,maxVar,learner,k)
  message <- paste("subject ", nameSubject , "using classifier KNN with k ",    k, " with 10x10 Fold CV for ressampling. The error rate in test is ", errorRate)
  print(message)

  
}

```
KNN function for all the methods
```{r}
subject.KNNMethod(subject1_raw,"1 with raw data")
subject.KNNMethod(subject2_raw,"2 with raw data")
subject.KNNMethod(subject3_raw,"3 with raw data")

subject.KNNMethod(subject1_psd.refactor,"1 with psd data")
subject.KNNMethod(subject2_psd.refactor,"2 with psd data")
subject.KNNMethod(subject3_psd.refactor,"3 with psd data")
```


Set up a diferent seed for non linear methods
```{r}
set.seed (4567)
```
Non-linear functions:
```{r}
subject.get.train<- function(number)
{
  if(number ==1){tr<- trainControl(method="LGOCV")}
  
  else if(number == 2)  {tr<- trainControl (method="repeatedcv", number=10, repeats=3)}  
  else if(number==3){tr<- trainControl(method="cv", number=2)} 

       
  return(tr)
}

subject.get.message.train<- function(number)
{

  if(number ==1){message<- "LGOCV"}
  else if(number == 2)  {message<- "repeatcv resampling 10 folds and 3 repeats"}   
  else if(number==3){message <- "cv resampling 10 folds"} 
       
  return(message)
}


subject.get.method <- function(number,subject,maxVar,trc)
{
  if(number == 1){method<-subject.get.method.resample.RBFKernel(subject,maxVar,trc)}
  else if(number==2){method<- subject.get.method.resample.MLP(subject,maxVar,trc)} 
  else if(number == 3){method<- subject.get.method.resample.RandomForest(subject,maxVar,trc)}
  else {method<- subject.get.method.resample.RBFNN(subject,maxVar,trc)}   
  return(method)
}




subject.get.message.method<- function(number)
{

  if(number==1){message <- "RBFKernel "} 
  else if(number == 2)  {message<- " MLP "}   
  else if(number ==3){message<- "Random Forest "}
  else {message<- "RBFNN "}
       
  return(message)
}

```
Non linear functions for printing messages
```{r}
subject.get.message.from.bestTune<- function(number,bestTune)
{

  if(number==1){message <- subject.get.message.RBFKernel.from.bestTune(bestTune)} 
  else if(number == 2)  {message<- subject.get.message.MLP.from.bestTune(bestTune)}   
  else if(number ==3){message<- subject.get.message.RandomForest.from.bestTune(bestTune)}
  else {message<- subject.get.message.RBFNN.from.bestTune(bestTune)}
       
  return(message)
}
subject.get.message.MLP.from.bestTune <- function(bestTune)
{
  message <- paste(" size : ",bestTune["size"]," and decay: ", bestTune["decay"])
  return(message)

    
}

subject.get.message.RBFNN.from.bestTune <- function(bestTune)
{
  message <- paste(" size : ",bestTune["size"])
  return(message)
    
  
}

subject.get.message.RandomForest.from.bestTune <- function(bestTune)
{
  message <- paste(" mtry : ",bestTune["mtry"])
  return(message)
  
}

subject.get.message.RBFKernel.from.bestTune <- function(bestTune)
{
  message <- paste(" C : ",bestTune["C"]," and sigma: ", bestTune["sigma"])
  return(message)
}


subject.resultsNN <- function(subject,subject.model,maxVar,subject.learn){
  ntrain <- length(subject.learn)
  ntest <-  dim(subject[-subject.learn,])[1]

  
  p2 <- as.factor(predict (subject.model, newdata=subject[-subject.learn,1:maxVar], type="raw"))
  subject.resultsTest <- table(p2,subject[-subject.learn,(maxVar+1)])
  subject.errorRate.Test <- 1-(sum(diag(subject.resultsTest))/ntest)
  return(subject.errorRate.Test)

}
```

```{r}
subject.get.method.resample.MLP<- function(subject,maxVar,trc)
{
  sizes <- 2*seq(1,10,by=1)
  decays <- c(0.0001,0.001,0.01,0.1,1)
  
  
  model.10x10CV <- caret::train (y=subject[,(maxVar+1)] ,x=  subject[,1:maxVar],method='nnet', maxit = 100, trace = FALSE,tuneGrid = expand.grid(.size=sizes,.decay=decays ),preProcess = c("center", "scale"), trControl=trc,allowParallel=TRUE)
  return(model.10x10CV)
    
}



subject.get.method.resample.RBFNN<- function(subject,maxVar,trc)
{
  sizes <- 2*seq(1,10,by=1)
  model.10x10CV <- caret::train (x= subject[,1:maxVar],y= subject[,(maxVar+1)], method='rbf', maxit = 100, trace = FALSE,preProcess = c("center", "scale"),tuneLength = 20, trControl=trc)
  return(model.10x10CV)
    
  
}

subject.get.method.resample.RandomForest<- function(subject,maxVar,trc)
{

  model.10x10CV <- caret::train (x= subject[,1:maxVar],y= subject[,(maxVar+1)],method='parRF',preProcess = c("center", "scale"), maxit = 100, trace = FALSE, trControl=trc,tuneLength=20,allowParallel=TRUE)
  return(model.10x10CV)
  
}

subject.get.method.resample.RBFKernel<- function(subject,maxVar,trc)
{
  model.10x10CV <- caret::train (x= subject[,1:maxVar],y= subject[,(maxVar+1)],method='svmRadial', preProcess = c("center", "scale"),maxit = 100, trace = FALSE,tuneLength = 20, trControl=trc)
  return(model.10x10CV)
  
}


subject.resampling <- function(subject,nameSubject)
{
 
  maxVar <- dim(subject)[2]-1
  subject[,(maxVar+1)] <- as.factor(subject[,(maxVar+1)])
  learner <- subject.get.learner(subject,maxVar)
  trainData <- subject[learner,]
  number <-  3
  tr <-subject.get.train(number)
  for( numberMethod in 1:3){
     subject.model <- subject.get.method(numberMethod,trainData,maxVar,tr)
     
     messageMethod <-subject.get.message.method(numberMethod)
       messagePlot <- paste("Plot between the \n estimates of performance \n and the tuning parameters  \n using the method ", messageMethod, " for Subject ", nameSubject)
     plt <- ggplot(subject.model) + ggtitle(messagePlot)
     print(plt)
     
     message.from.BestTune<- subject.get.message.from.bestTune(numberMethod,subject.model$bestTune)
     subject.error.TrainTest <-subject.resultsNN(subject,subject.model,maxVar,learner) 
     
     message <- paste(messageMethod, " with 10foldCV for sampling, the best parameter are ", message.from.BestTune, " with TrainTest resampling ", subject.error.TrainTest," for Subject ", nameSubject)
     print(message)

    }
}
```



```{r}
subject.resampling(subject3_psd.refactor,"3 using psd data")
subject.resampling(subject2_psd.refactor,"2 using psd data")
subject.resampling(subject1_psd.refactor,"1 using psd data")

subject.resampling(subject3_raw,"3 using raw data")
subject.resampling(subject2_raw,"2 using raw data")
subject.resampling(subject1_raw,"1 using raw data")
```